Distributed Decision Making for Dynamic Formation of Web Services Communities

### 2015-05-09 papers/webservice-coalition-learning-omar-template/elsarticle-template-num.tex
rabeb comments

\begin{abstract}
Web services are loosely-coupled business applications willing to cooperate in distributed settings within different groups called communities. Communities aim to provide web services with better visibility, efficiency, market share and total payoff. A number of mechanisms and models have been recently proposed to aggregate web
services and make them cooperate within their communities. However, existing community models mostly focus on short term goals. Moreover, these models generally suffer from high computational complexity, which makes decision making about community formation in real-time extremely hard, time consuming, and even impossible in some complex cases. In this paper, we propose a strategic distributed decision making mechanism called DDM that regulates web service agents' decision making process, which is about deciding with whom to cooperate. This model first generates a trained set of data based on information obtained from large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The decision making mechanism implements a decision tree of possible viable strategies and their long-term expected utility gain.
%Utilizing the training model, the decision making mechanism implements a decision tree with an estimation of long-term utility gain for the web services. Communities can used the trained model and instantly choose best-response strategies considering their long-term gain. 
Our model helps 1) communities of web services efficiently find the appropriate web service to invite as new member; and 2) single web services find best communities to join. After being trained, web services get to compute expectations as utilities they would gain if cooperating with communities of different characteristics. In contrast, communities can consider the training into account and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated initiated by a web service. In general, DDM equips web services with efficient methods of foreseeing how their choices of actions would impact their short-term and long-term goals, therefore opting for better decisions. We performed experiments using real data from 142 users on 4,532 web services on 64 different time slots. The experimental results, show that our algorithms provide web services and community owners, in real-world-like environments, with applicable and near-perfect decision making mechanisms.
\end{abstract}
\begin{IEEEkeywords}
Web services; Community of services; Distributed Decision Making;
\end{IEEEkeywords}

@@

\IEEEPARstart{O}{ver} the past years, online services have become an important part of many scalable business applications. The increasing reliance on online service providers has significantly influenced the way web services are engineered. Therefore, given the dynamic and unpredictable nature of the Internet, delivering high quality services is still a critical and challenging issue. One practical solution towards delivering high quality services is utilizing intelligent decision making agents. These agents aim at maximizing their gain by exploring the best ways to provide services that satisfy end users\cite{Zeng:2003:QDW:775152.775211, 10.1109/ARES.2008.7, Demirkan2013412, journals/tsc/ZhengZYB13, Josang:2007:STR:1225318.1225716}. However, agent-based web services are functionally limited in the sense that they cannot handle a large number of requests at the same time without compromising the quality of service provided. Recent developments have attempted to shift web services from passive models, consisting of individual components, to models made up of autonomous and group-based components that share common goals. In group-based models, interaction, composition, and cooperation are the key challenges \cite{ICWS2011-1, SCC2011-1, journals/mags/BaldoniBM10, journals/jcss/CasadoYT13} that directly impact the group's overall performance in achieving common goals. To that end, we see the emergence of web service \emph{communities}, which consist of services with similar functionalities but distinct nonfunctional properties \cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038, Medjahed05adynamic, 10.1109/ARES.2008.7}. A \emph{community} of web services runs continuous performance assessment functions that regulate web services' interactions and manage their composition and cooperation.

The advantages of community of web services are that the community facilitates the discovery of web services and provides better quality of service compared to passive models with individual components. Communities act as abstract web services, communicating with external entities via the same standard protocols that a normal web service employs. But, they regulate the service process via sophisticated internal communication protocols, thereby providing services based on the combined work of a number of web services. The downside to communities is the complexity of management involved in finding and inviting individual web services and managing the overall quality of the combined work of several web services.
% because although they have similar functionality, they have different attitudes. 
When interacting with communities of web services, users send their requests to the coordinators of the community, which play the role of community representative or access point. The community coordinator is responsible of receiving tasks and delivering services. Moreover, as community representative, it verifies the credentials of new web services before accepting them into the community and kicks web services that could harm the value of the community.

\textbf{Challenge and Problem Statement.} In recent work, communities of web services have been proposed in order to facilitate discovery of web services, improve the Quality of Service (QoS), and help individual services find better market share and opportunities. However, there is a twofold challenge that yet to be addressed. The first part is community development, which involves finding the right web services for the community, satisfying the objectives of all the web services and community coordinators involved within the community; and the second part is from the individual web service's side because each web service aims to join the community that would guarantee the best outcome. The advocated solutions \cite{conf/webist/MaamarLBTS07, 10.1109/ARES.2008.7, journals/soca/XuYLZB11, 10.1109/TSC.2012.12, managing-hela-jalel, DBLP:conf/IEEEscc/KhosravifarABT11, DBLP:conf/IEEEscc/LimTMB12} have attempted to address these research problems. These research problems are reflected by the two challenge parts: 
\begin{enumerate}
	\item The solutions consider the architecture of centralized management for communities where most of the decisions are made by the centralized coordinator. The problem is that in real world scenarios, decisions made by independent service providers are highly distributed.
	\item The solutions either propose complex algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to find the optimal strategy to follow, or oversimplify the problem by eliminating important parameters and using approximation techniques to make the algorithms tractable \cite{10.1109/TSC.2012.12}.
\end{enumerate}
These approximation methods sometimes negatively influence the outcome because simplifying the constraints may cause important aspects of the problem to be ignored.

@@

\textbf{Contributions.} In this paper, we introduce DDM, a Distributed Decision Making model for  community formation that regulates web service agents’ decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. The decision model in extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and how to select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree is created from the past data and reflects comprehensive vision about agents' attitude of action selection. However, the tree is subject to get continuously modified to adapt with environmental changes.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems. 
The decision model provides web services with enough information which helps those services to efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in decision making process and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. In other words, the web services agents are equipped with a recommender system to suggest beneficial action. Moreover, communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact their short-term and long-term goals; therefore, opting for best decision available.

@@

\section{Preliminaries and Challenging Issues}\label{s:preliminaries}
%In this section, we first present the architecture of DDM. We explore the characteristics of intelligent service agents and the features we extract for training. To do this, we first discuss some preliminaries.
In this section, we discuss the preliminary concepts of communities of web service and we introduce the challenges of community formation.

@@

The World Wide Web Consortium (W3C) defines web services as ``software systems designed to support interpretable machine-to-machine interaction over a network. It has an interface described in a machine-processable format (specifically WSDL). Other systems interact with the web service in a manner prescribed by its description using SOAP messages, typically conveyed using HTTP with XML serialization in conjunction with other Web-related standards.'' When developers declare a new web service, it will be
discovered based on its description, which fully discloses its functionalities. Developers also have to declare a public interface and a readable documentation to help other developers when integrating different services \cite{w3cwsdl}. Nowadays, web API, standards that do not require XML-based web service protocols like SOAP and WSDL are also emerging. They are called REST (representational state transfer) services, which are moving towards simpler communication protocols. 
%They are not restricted 
%to XML formats, recently JSON, a human readable and simpler format 
%is becoming popular among online service providers.

@@

In this paper, we abstract web services as rational agents\footnote{The term rational is used here in the sense that web services are utility maximizers} that provide services to end users. They aim to maximize their individual income  by receiving enough requests from end users. In order to increase their revenues, web services seek for more tasks if they have the capacity and throughput to do so. Web services can join communities to enhance efficiency by collaborating with others, to have access to broad market share, and for the opportunity to receive a bigger task pool from end users. Furthermore, the high reliance on web services has resulted in increased quality expectations from end users. Communities of web services can provide higher availability, performance, reliability, and recovery for end users.

@@

The community of web services is essentially a virtual group of web services having similar and complementary functionalities\cite{DBLP:journals/ijebr/MaamarSTBB09}. Communities aggregate web services and communicate with other entities such as UDDI registries and users, using identical protocols to those used by single web services. Web services join communities to increase utility by having a larger market share and task pool. Community coordinators are responsible for community development, managing membership requests from web services and distributing user tasks among the community members. Community coordinators try to attract quality web services to join and keep the community as stable and productive as possible to gain a better reputation and user satisfaction, which increases the community's market share. How web services reside within communities and how communities of web services are engineered is described comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}.

\subsection{The Join Challenge}\label{s:tjc}
Sometimes, web services can increase their overall utility by collaborating with other web services as communities. This collaboration helps them with better ways of sharing resources and having higher reputation, better market share and wider visibility. Web services and communities come with different internal and external parameters, and the long-term outcome is dependent on these parameters.

@@

	\begin{equation}
		 Th_{C} = \left\{ 
			\begin{array}{l l}
				Th_{w} & \quad \text{if $|C| = 1$}\\
				\sum_{w \in C}{(Th_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

@@

	\begin{equation}
		A_{C} = \left\{ 
			\begin{array}{l l}
				A_{w} & \quad \text{if $|C| = 1$}\\
				1-\prod_{w \in C}{(1-A_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

@@

	\begin{equation}
		Et_{C} = \left\{ 
			\begin{array}{l l}
				Et_{w} & \quad \text{if $|C| = 1$}\\
				max_{w \in C}{(Et_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

@@

	We normalize the range of these features so that each feature contributes approximately proportionately to the final utility outcome value. We adopt the \emph{standardization} method. We subtract the \emph{mean} from each feature then we divide the values of each feature by its \emph{standard deviation}.

@@

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} is to other web services. It is the difference between a web service's \emph{execution time} metric and the maximum value of execution time of all the other web services, about which the web service has information of them. The less value means web service has better execution time compared to other web services.
	\begin{equation}\label{exp_1:f}
		exp_1 = Et_{i} - Et_{max}
	\end{equation}
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between a web service's \emph{throughput} metric and the maximum value of throughput of all the other web services,, about which the web service has information of them. The less value means web service has better throughput compared to other web services.
	\begin{equation}\label{exp_2Lf}
		exp_2 = Th_{max} - Th_{i}
	\end{equation}
\end{itemize}

A community of web services is abstracted as a standard web service that provides services that share the same internal and external features mentioned for web services. Inside, they adopt different protocols to manage the community; however, other web services or end users can communicate with communities of web services, using standard protocols, and inner details will be encapsulated for their clients.

@@

Communities of web services usually employ an implementation of Contract-Net protocol for task distribution, in which services opt for incoming tasks, and receive the tasks for which they ask. However, in our model, our community members would try to distribute tasks based on their capabilities and the QoS parameters provided by the web services. We have used a slightly modified \emph{weighted fair queuing} method to distribute tasks among community members. The goal is to allocate incoming tasks to web services with a rate matching the throughput value of $Th_{ws}$. In the \emph{weighted fair queuing} method, the input flow is multiplexed along different paths; however, in our case, if the rate of incoming tasks is less than the community's total throughput $(Th_{C})$, which is the summation of throughput values of the web services in the community, some of the input tasks will be queued and served with a delay. 
%Thus, the amount of tasks performed by the community is $\sum_{ws}{Th_{ws}}$ when $\sum_{ws}{Th_{ws}} \leq R_{C}$. 
However, when the incoming task rate is less than the throughput of the community, the \emph{weighted fair queuing} algorithm assigns a weighted task rate of $|input Task Rate| \times \frac{Th_{ws}}{\sum_{ws}{Th_{ws}}}$ for each web service ($ws$) within the community.

While distributing tasks, the community members can verify the performance, throughput and quality of service of tasks being performed by web services. The community can assess if those web services are capable of performing the number of tasks they advertised. If for any reason, there is a decline in the quality metric or throughput, the community can consider the new values as a benchmark for future performance calculations, and penalize the suspicious web services. This way, players will have incentive to truthfully disclose their actual capabilities in order to maximize profit from the community and to avoid being penalized. In addition, the system should be dynamic enough to detect and react to web services' quality metrics variation, as over time, web service metrics may degrade or improve, changes to which the community should adjust.
% Therefore its easy for the system to encourage players to be in some sense incentive compatible in the way that they would profit best by truthfully revealing their capabilities. Also it is important to be dynamic enough to consider web services which may have their quality metrics degraded or even improved over time for any reason and be able to adjust the community with new parameters.

@@

%We used a data set extracted from running web services. In this data set, 
Each web service is associated with a number of features that reflect its functionality. These web services operate in an online environment and are continuously assigned tasks to handle. Sometimes, they pass tasks around, distributing sub-tasks among themselves and handling tasks in a collaborative effort. To engage in communities, we add additional features that reflect web services' utility as a result of joining other web services to form a community. The additional features are computed based on two assumptions that we adopt for a community to be formed. Our raw data set provides us with 64 different time slots of extracted features for each web service. Using this data we can build a synthetic data set that contains features of a large number of web services in a sequence of time. Therefore, we train a learning model that adopts the trend of joining a community and use the model to predict/find the appropriate community for other web services. To train our model, we build a decision tree as a benchmark for our decision making mechanism. The decision tree is built by training the real data obtained from operating web services and extracting features related to their performance, either alone or as part of a joint effort with other web services. As a result, we are able to classify web services as either single or joined and propose the communities that are best fitted to the web services' requirements.

@@

We have adopted the web service data set provided by \cite{10.1109/ISSRE.2011.17}. The raw data provides real-world QoS evaluation results from several users on 5,825 web services over 64 different time frames\footnote{http://www.wsdream.net/}. In this data set, each web service is associated with a number of features that reflect its functionality. By summarizing the data provided for each web service over different time slots, we have acquired three quality features for each individual web service. Their quality features are the three internal features introduced in Section \ref{s:if}: \emph{throughput}, \emph{availability} and \emph{execution time}. 
%To engage in communities, we add two additional external features that reflect web services estimation of its quality metrics compared to other web services. 
Therefore, we have generated an array of web services with three distinct features.

@@

\begin{table*}[ht]
\caption{An example of $gain$ matrix for 3 different communities and their combinations} % title of Table
\centering % used for centering table
{\renewcommand{\arraystretch}{1.2}
\begin{tabular}{c|c c c c c c} % centered columns (4 columns)
\hline\hline %inserts double horizontal lines
 & \textless348\textgreater & \textless1934\textgreater & \textless2117\textgreater & \textless348, 1934\textgreater & \textless1934, 2117\textgreater & \textless348, 1934, 2117\textgreater \\ [0.5ex] % inserts table
%heading
\hline % inserts single horizontal line
\textless348\textgreater & - & 0.282708 & 1.027081 & 0.282708 & 18.027081 & 18.027081 \\
\textless1934\textgreater & -2.637483 & - & 6.969072 & -2.637483 & 5.509583 & 4.387725 \\
\textless2117\textgreater & 5.027081 & 2.969072 & - & 5.509583 & 2.969072 & 5.509583 \\
\textless348, 1934\textgreater & 0.0 & 0.0 & -3.851432 & - & -3.851432 & -3.851432 \\
\textless1934, 2117\textgreater & 2.969072 & 0.0 & 0.0 & 2.969072 & - & 2.969072 \\
\textless348, 1934, 2117\textgreater & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & - \\ [1ex] % [1ex] adds vertical space
\hline %inserts single line
\end{tabular}
}
\label{table:nonlin} % is used to refer this table in the text
\end{table*}

@@

Let $P_{C_i}^t = \{C_1 >_{i}^t C_2 >_{i}^t .... >_{i}^t C_n\}$ be an ordered set of preferences for community $C_i$ at time $t$. Based on this ordered set, we define $K^t(C_i, k)$ as a set of the $k$ most preferred communities of community $i$ at time $t$.
%\subsubsection{feature vector generation}\label{sss:fvg}
\begin{equation}\label{h_t_pref_top}
\begin{split}				
K^t(C_i, k) = &\Big\{C_x | C_x >_{i}^t C_y; \forall C_y \in CFVS\ \\
				      &\wedge C_y \notin \{K^t(i, 1),K^t(i, 2),...,K^t(i, k-1)\} \Big\}				
\end{split}
\end{equation}
Based on $K^t(C_i, k)$, we define a set of communities for $C_i$ in which they are the k most preferred communities for community $i$ and also community $i$ belongs to the k most preferred communities out of all of the communities in set $K^t(C_i, k)$. This basically implies the preference is both-sided.
\begin{equation}\label{l_t_top_both}
\begin{split}	
L^t(C_i,k) = &\Big\{C_j | C_j \in K^t(C_i, k) \\
             &\wedge C_i \in \bigcup_{C_x \in K^t(C_j, k)}K^t(C_x, k)\Big\}
\end{split}
\end{equation}

@@

Figure \ref{stats3} depicts the average community size over time, which essentially represents the number of new communities being formed. Once again, these results show that the runs with higher search rates tend to find appropriate communities to join faster.

>>>>

\begin{abstract}
Web services are loosely-coupled business applications willing to cooperate in distributed settings within different groups called communities. Communities aim to provide web services with better visibility, efficiency, market share and total payoff. A number of mechanisms and models have been recently proposed to aggregate web
services and make them cooperate within their communities. However, existing community models mostly focus on short term goals. Moreover, these models generally suffer from high computational complexity, which makes decision making about community formation in real-time extremely hard, time consuming, and even impossible in some complex cases. In this paper, we propose a strategic distributed decision making mechanism called DDM that regulates web service agents' decision making process, which is about deciding with whom to cooperate. This model first generates a trained set of data based on information obtained from large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The decision making mechanism implements a decision tree of possible viable strategies and their long-term expected utility gain.
%Utilizing the training model, the decision making mechanism implements a decision tree with an estimation of long-term utility gain for the web services. Communities can used the trained model and instantly choose best-response strategies considering their long-term gain. 
Our model helps 1) communities of web services efficiently find the appropriate web service to invite as new member; and 2) single web services find best communities to join. After being trained, web services get to compute expectations as utilities they would gain if cooperating with communities of different characteristics. In contrast, communities can consider the training into account and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated by a web service. In general, DDM equips web services with efficient methods of foreseeing how their choices of actions would impact their short-term and long-term goals, therefore opting for better decisions. We performed experiments using real data from 142 users on 4,532 web services on 64 different time slots. The experimental results, show that our algorithms provide web services and community owners, in real-world-like environments, with applicable and near-optimal decision making mechanisms.
\end{abstract}
\begin{IEEEkeywords}
Web services; Community of services; Distributed Decision Making;
\end{IEEEkeywords}

@@

\IEEEPARstart{O}{ver} the past years, online services have become an important part of many scalable business applications. The increasing reliance on online service providers has significantly influenced the way web services are engineered. Therefore, given the dynamic and unpredictable nature of the Internet, delivering high quality services is still a critical and challenging issue. One practical solution towards delivering high quality services is utilizing intelligent decision making agents. These agents aim at maximizing their gain by exploring the best ways to provide services that satisfy end users\cite{Zeng:2003:QDW:775152.775211, 10.1109/ARES.2008.7, Demirkan2013412, journals/tsc/ZhengZYB13, Josang:2007:STR:1225318.1225716}. However, agent-based web services are functionally limited in the sense that they cannot handle a large number of requests at the same time without compromising the quality of service provided. Recent developments have attempted to shift web services from passive models, consisting of individual components, to models made up of autonomous and group-based components that share common goals. In group-based models, interaction, composition, and cooperation are the key challenges \cite{ICWS2011-1, SCC2011-1, journals/mags/BaldoniBM10, journals/jcss/CasadoYT13} that directly impact the group's overall performance in achieving common goals. To that end, we see the emergence of web service \emph{communities}, which consist of grouping of services with similar functionalities but distinct nonfunctional properties \cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038, Medjahed05adynamic, 10.1109/ARES.2008.7}. A \emph{community} of web services runs continuous performance assessment functions that regulate web services' interactions and manage their composition and cooperation.

Web Service communities have the advantages of facilitating web service discovery and providing better quality of service compared to individual components. Communities act as abstract web services, communicating with external entities via the same standard protocols that a normal web service employs. But, they regulate the service process via sophisticated internal communication protocols, thereby providing services based on the combined work of a number of web services. The downside to communities is the complexity of management involved in finding and inviting individual web services and managing the overall quality of the combined work of several web services.
% because although they have similar functionality, they have different attitudes. 
When interacting with communities of web services, users send their requests to the coordinators of the community, which play the role of community representative or access point. The community coordinator is responsible of receiving tasks and delivering services. Moreover, as community representative, it verifies the credentials of new web services before accepting them into the community and kicks web services that could harm the value of the community.

\textbf{Challenge and Problem Statement.} In recent work, communities of web services have been proposed in order to facilitate discovery of web services, improve the Quality of Service (QoS), and help individual services find better market share and opportunities\cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038, Medjahed05adynamic, 10.1109/ARES.2008.7}. However,  two important challenges are to be addressed: 1) choice of best web services during community development from the community perspective and 2) choice of the best community to join from the web service perspective. The advocated solutions \cite{conf/webist/MaamarLBTS07, 10.1109/ARES.2008.7, journals/soca/XuYLZB11, 10.1109/TSC.2012.12, managing-hela-jalel, DBLP:conf/IEEEscc/KhosravifarABT11, DBLP:conf/IEEEscc/LimTMB12} have attempted to address these research problems. However the solutions have not addressed these two challenging issues: 
\begin{enumerate}
	\item The solutions consider the architecture of centralized management for communities where most of the decisions are made by the centralized coordinator. The problem is that in real world scenarios, decisions made by independent service providers are highly distributed.
	\item The solutions either propose complex algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to find the optimal strategy to follow, or oversimplify the problem by eliminating important parameters and using approximation techniques to make the algorithms tractable \cite{10.1109/TSC.2012.12}.
\end{enumerate}
These approximation methods sometimes negatively influence the outcome because simplifying the constraints may negatively cause important aspects of the problem to be ignored.

@@

\textbf{Contributions.} In this paper, we introduce DDM, a Distributed Decision Making model for  community formation that regulates web service agents’ decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. The decision model in extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and how to select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree which is initially created from the past data and reflects comprehensive vision about agents' attitude of action selection based on agents' past expiriences. However, the tree based on recieving new feedback and outcome from actions, is subject to get continuously modified to adapt with environmental changes.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems. 
The decision model provides web services with enough information which helps those services to efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in decision making process and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. In other words, the web services agents are equipped with a recommender system to suggest beneficial action. Moreover, communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact their short-term and long-term goals; therefore, opting for best decision available.

@@

\section{Preliminaries and Challenges in WS Communities}\label{s:preliminaries}
%In this section, we first present the architecture of DDM. We explore the characteristics of intelligent service agents and the features we extract for training. To do this, we first discuss some preliminaries.
In this section, we discuss the preliminary concepts of communities of web service and we introduce the challenges of community formation.

@@

The World Wide Web Consortium (W3C) defines web services as ``software systems designed to support interpretable machine-to-machine interaction over a network. It has an interface described in a machine-processable format (specifically WSDL). Other systems interact with the web service in a manner prescribed by its description using SOAP messages, typically conveyed using HTTP with XML serialization in conjunction with other Web-related standards.'' When developers declare a new web service, it will be
discovered based on its description, which fully discloses its functionalities. Developers also have to declare a public interface and a readable documentation to help other developers when integrating different services \cite{w3cwsdl}. Nowadays, web API, standards that do not require XML-based web service protocols like SOAP and WSDL are also emerging. They are called RESTful (representational state transfer) services, which are moving towards simpler communication protocols. 
%They are not restricted 
%to XML formats, recently JSON, a human readable and simpler format 
%is becoming popular among online service providers.

@@

In this paper, we abstract web services as rational agents\footnote{The term rational is used here in the sense that web services are utility maximizers} that provide services to end users. They aim to maximize their individual income  by receiving enough requests from end users. In order to increase their revenues, web services seek for more tasks if they have the capacity and throughput to do so. Web services can join communities to enhance efficiency by collaborating with others, to have access to broad market share, and for the opportunity to receive a bigger task pool from end users. 
%Furthermore, the high reliance on web services has resulted in increased quality expectations from end users. Communities of web services can provide higher availability, performance, reliability, and recovery for end users.

@@

The community of web services is essentially a virtual group of web services having similar functionalities\cite{DBLP:journals/ijebr/MaamarSTBB09}. Communities aggregate web services and communicate with other entities such as UDDI registries and users, using identical protocols to those used by single web services. Web services join communities to increase utility by having a larger market share and task pool. Community coordinators are responsible for community development, managing membership requests from web services and distributing user tasks among the community members. Community coordinators try to attract quality web services to join and keep the community as stable and productive as possible to gain a better reputation and user satisfaction, which increases the community's market share. How web services reside within communities and how communities of web services are engineered is described comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}.

\subsection{The Join Challenge}\label{s:tjc}
Sometimes, web services can increase their overall utility by collaborating with other web services as communities. This collaboration helps them with better ways of sharing resources and having higher reputation, better market share and wider visibility. Web services and communities come with different quality metric parameters, and the long-term outcome is dependent on these parameters.

@@

	\begin{equation}
		 Th_{C} = \sum_{w \in C}{(Th_{w})}
	\end{equation}

@@

	\begin{equation}
		A_{C} = 1-\prod_{w \in C}{(1-A_{w})} 
	\end{equation}

@@

	\begin{equation}
		Et_{C} = max_{w \in C}{(Et_{w})}
	\end{equation}

@@

	We normalize the range of these features so that each feature contributes proportionally to the final utility outcome value. We adopt the \emph{standardization} method. We subtract the \emph{mean} from each feature then we divide the values of each feature by its \emph{standard deviation}.

@@

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} is to other web services. It is the difference between a web service's \emph{execution time} metric and the maximum value of execution time of all the other web services. The smaller value means web service has better execution time compared to other web services.
	\begin{equation}\label{exp_1:f}
		exp_1 = Et_{i} - Et_{max}
	\end{equation}
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between a web service's \emph{throughput} metric and the maximum value of throughput of all the other web services. The less value means web service has better throughput compared to other web services.
	\begin{equation}\label{exp_2Lf}
		exp_2 = Th_{max} - Th_{i}
	\end{equation}
\end{itemize}

A community of web services is abstracted as a standard web service that provides services that share the same internal and external features mentioned above for the web services. Inside, they adopt different protocols to manage the community; however, other web services or end users can communicate with communities of web services, using standard web service protocols.

@@

Communities of web services usually employ an implementation of Contract-Net protocol for task distribution, in which services opt for incoming tasks, and receive the tasks for which they ask. However, in our model, our community members would try to distribute tasks based on their capabilities and the QoS parameters provided by the web services. We have used a slightly modified \emph{weighted fair queuing} method to distribute tasks among community members. The goal is to allocate incoming tasks to web services with a rate matching the throughput value of $Th_{ws}$. In the \emph{weighted fair queuing} method, the input flow is multiplexed along different paths; however, in our case, if the rate of incoming tasks is less than the community's total throughput $(Th_{C})$, which is the summation of throughput values of the web services in the community, some of the input tasks will be queued and served with a delay. 
%Thus, the amount of tasks performed by the community is $\sum_{ws}{Th_{ws}}$ when $\sum_{ws}{Th_{ws}} \leq R_{C}$. 
When the incoming task rate is less than the throughput of the community, the \emph{weighted fair queuing} algorithm assigns a weighted task rate of $|input Task Rate| \times \frac{Th_{ws}}{\sum_{ws}{Th_{ws}}}$ for each web service ($ws$) within the community.

While distributing tasks, the community can verify the performance, throughput and quality of service of tasks being performed by web services. The community can assess if those web services are capable of performing the number of tasks they advertised. If for any reason, there is a decline in the quality metric or throughput, the community can consider the new values as a benchmark for future performance calculations, and penalize the suspicious web services. This way, players will have incentive to truthfully disclose their actual capabilities in order to maximize profit from the community and to avoid being penalized. In addition, the system should be dynamic enough to detect and react to web services' quality metrics variation, as over time, web service metrics may degrade or improve, changes to which the community should adjust.
% Therefore its easy for the system to encourage players to be in some sense incentive compatible in the way that they would profit best by truthfully revealing their capabilities. Also it is important to be dynamic enough to consider web services which may have their quality metrics degraded or even improved over time for any reason and be able to adjust the community with new parameters.

@@

%We used a data set extracted from running web services. In this data set, 
Each web service is associated with a number of features that reflect its functionality. These web services operate in an online environment and are continuously assigned tasks to handle. To engage in communities, we add additional features that reflect web services' utility as a result of joining other web services to form a community. The additional features are computed based on two assumptions that we adopt for a community to be formed. We have adopted the web service data set provided by \cite{10.1109/ISSRE.2011.17}. The raw data provides real-world QoS evaluation results from several users on 5,825 web services over 64 different time frames\footnote{http://www.wsdream.net/}. In this data set, each web service is associated with a number of features that reflect its functionality. 
%Our raw data set provides us with 64 different time slots of extracted features for each web service. 
Using this data we can build a synthetic data set that contains features of a large number of web services in a sequence of time. Therefore, we train a learning model that adopts the trend of joining a community and use the model to predict/find the appropriate community for other web services. To train our model, we build a decision tree as a benchmark for our decision making mechanism. The decision tree is built by training the real data obtained from operating web services and extracting features related to their performance, either alone or as part of a joint effort with other web services. As a result, we are able to classify web services as either single or joined and propose the communities that are best fitted to the web services' requirements.

@@

By summarizing the data provided for each web service over different time slots, we have acquired three quality features for each individual web service. Their quality features are the three internal features introduced in Section \ref{s:if}: \emph{throughput}, \emph{availability} and \emph{execution time}. 
%To engage in communities, we add two additional external features that reflect web services estimation of its quality metrics compared to other web services. 
Therefore, we have generated an array of web services with three distinct features.

@@

\begin{table*}[ht]
\caption{An example of $gain$ matrix for 3 different communities and their combinations} % title of Table
\centering % used for centering table
{\renewcommand{\arraystretch}{1.2}
\begin{tabular}{c|c c c c c c} % centered columns (4 columns)
\hline\hline %inserts double horizontal lines
 & \textless348\textgreater & \textless1934\textgreater & \textless2117\textgreater & \textless348, 1934\textgreater & \textless1934, 2117\textgreater & \textless348, 1934, 2117\textgreater \\ [0.5ex] % inserts table
%heading
\hline % inserts single horizontal line
\textless348\textgreater & - & 0.282708 & 1.027081 & 0.282708 & 18.027081 & 18.027081 \\
\textless1934\textgreater & -2.637483 & - & 6.969072 & -2.637483 & 5.509583 & 4.387725 \\
\textless2117\textgreater & 5.027081 & 2.969072 & - & 5.509583 & 2.969072 & 5.509583 \\
\textless348, 1934\textgreater & - & - & -3.851432 & - & -3.851432 & -3.851432 \\
\textless1934, 2117\textgreater & 2.969072 & - & - & 2.969072 & - & 2.969072 \\
\textless348, 1934, 2117\textgreater & - & - & - & - & - & - \\ [1ex] % [1ex] adds vertical space
\hline %inserts single line
\end{tabular}
}
\label{table:nonlin} % is used to refer this table in the text
\end{table*}

@@

Let $P_{C_i}^t = \{C_1 >_{i}^t C_2 >_{i}^t .... >_{i}^t C_n\}$ be an ordered set of preferences for community $C_i$ at time $t$. Based on this ordered set, we define $K^t(C_i, k)$ as a set of the $k$ most preferred communities of community $i$ at time $t$.
%\subsubsection{feature vector generation}\label{sss:fvg}
\begin{equation}\label{h_t_pref_top}
\begin{split}				
K^t(C_i, k) = &\Big\{C_x | C_x >_{i}^t C_y; \forall C_y \in CFVS\ \\
				      &\wedge C_y \notin \{K^t(C_i, 1),K^t(C_i, 2),...,K^t(C_i, k-1)\} \Big\}				
\end{split}
\end{equation}
Based on $K^t(C_i, k)$, we define a set of communities for $C_i$ in which they are the k most preferred communities for community $i$ and also community $i$ belongs to the k most preferred communities out of all of the communities in set $K^t(C_i, k)$. This basically implies the preference is both-sided.
\begin{equation}\label{l_t_top_both}
\begin{split}	
L^t(C_i,k) = &\Big\{C_j | C_j \in K^t(C_i, k) \\
             &\wedge C_i \in \bigcup_{C_x \in K^t(C_j, k)}K^t(C_x, k)\Big\}
\end{split}
\end{equation}

@@

Figure \ref{stats3} depicts the average community size over time, which essentially represents the number of new communities being formed. The results show once again the communities with higher search rates grow faster in size implying that the communities find appropriate web services to join with faster.
